---
title: "Refugee Resettlements in the USA"
author: "Ji In Choi (jic2124), Jung Ah Shin (js5569), Olivia Wang (yw3324), Tiffany Zhu (tz2196)"
date: "December 12, 2019"
output:
  html_document:
    toc: true
    toc_float: true
    theme: journal
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(readxl)
library(tidyverse)
library(plotly)
library(knitr)
```

# Introduction
According to the UNHCR, refugees are those who have been forced to leave their country due to violence, war, or persecution based on their race, religion, nationality, political opinion or particular social group. 
Furthermore, there are currently 25.9 million refugees in the world, indicating the dramatic growth in refugees over the past decade. This led us to question what the refugee resettlement trend has been for the past decade, and delve deeper than just the changes in the numbers of refugees. 

The United States is one of the nations that have one of the largest resettlement of refugees. However, in 2017, the US resettled the fewest number of refugees compared to the rest of the world.We are interested in answering the following questions to gain a better understanding of the refugee resettlements in the United States:

i. What insights can we gain from temporal exploratory data analysis of refugee settlement patterns in the US? Have there been increases/decreases in refugee settlements from 2009 - 2018?
ii. What insights can we gain from geographical visualization of refugee settlement patterns in the US? Why might some states have larger refugee settlements than others?
iii. What changes in demographic patterns (i.e., religion, gender, age, etc.) within the refugee population can we visualize? Can we observe any relationship between certain demographics and refugee settlements?


# Data Sources

We first collected data from RPC (Refugee Processing Center), that provides refugee arrival information by state and nationality, by destination and nationality, by nationality and religion, and by demographic profile. 

We can select the time frame, nationality. 
Since the RPC website does not allow for faceting by year, we had to download the files year by year, and clean the data into the format we want for data analysis.


# Data Transformation
## Cleaning the Data in R

1. From the website, we were able to download '.xlsx' files. Raw files can be found [here](https://github.com/tlzhu19/edav-final-project-41/tree/master/data/raw).
2. Wrote two functions to clean Excel sheet for a given year. 
    i. `clean_arrival` to clean the Excel files for all refugee resettlements for each state.
    ii. `clean_demographics` to clean the Excel files for demographic information for refugees from specific countries (namely Bhutan, Burma, DRC, Iraq, and Somalia).
3. Wrote another function, `combine_files`, to combine each year's Excel file into one.
4. Saved these as csv files and uploaded to GitHub to easily access.

```{r}
# Used to clean the Arrivals data
clean_arrival <- function(file, sheet_name) {
  dat <- read_excel(file, sheet = sheet_name)
  dat1 <- subset(dat, select = c(...6, ...8,...9))
  dat2 <-subset(dat, select = c(...15,...18,...19)) 
  dat3 <-subset(dat, select = c(...23,...25,...27))

  dat1<-dat1 %>% drop_na()
  dat2<-dat2 %>% drop_na()
  dat3<-dat3 %>% drop_na()

  # set first row (State, Cases, Inds) as column name
  colnames(dat1) <- as.character(unlist(dat1[1,]))
  dat1 = dat1[-1, ]
  colnames(dat2) <- as.character(unlist(dat2[1,]))
  dat2 = dat2[-1, ]
  colnames(dat3) <- as.character(unlist(dat3[1,]))
  dat3 = dat3[-1, ]

  # Grouping columns together
  combined <- rbind(dat1,dat2)
  combined <- rbind(combined,dat3)

  # State,Cases,Inds

  combined$Cases <-as.numeric(as.character(combined$Cases))
  combined$Inds <-as.numeric(as.character(combined$Inds))
  
  return (combined)
}


```

```{r}
# Used to clean the Demographics data
clean_demographics <- function(file, sheet_name) {
  df <- read_excel(file, sheet = sheet_name)
  
  if (sheet_name == 'Age Group') {
    column_names = c('Department of State','...2', '...3', '...4')
  }
  else if (sheet_name == 'Religion') {
    column_names = c('Department of State','...2', '...4', '...6')
  }
  else if (sheet_name == 'Ethnicity' | sheet_name == 'Education' | sheet_name == 'Native Language') {
    column_names = c('Department of State','...2', '...3', '...5')
  }
  
  df2 <- subset(df, select = column_names)
  df3 <- df2[complete.cases(df2), ]
  df4 = df3[-nrow(df3), ]
  
  #print(file)
  #file_replaced <- gsub(pattern="demographics/", replacement=".", file, fixed = TRUE)
  
  # file_unpacked <- unlist(strsplit(file_replaced, split='.', fixed=TRUE))
  file_unpacked <- unlist(strsplit(file, 'demographics/'))
  file_unpacked2 <- unlist(strsplit(file_unpacked[2], '_Demographics'))
  country_name = file_unpacked2[1]

  colnames(df4) = c(sheet_name, 'Male', 'Female', 'Total')
  
  df4$country <- rep(country_name, nrow(df4))
  
  return(df4)
}

```

```{r}
# Cleans and combines all the files in the directory given by path
combine_files <- function(path, data_cleaner, sheet_name, title) {
  df = data.frame()
  files <- list.files(path=path, pattern="*.xlsx", full.names=TRUE, recursive=FALSE)
  for (file in files){
    cleaned_data = data_cleaner(file, sheet_name)
    year_extract <- unlist(strsplit(file, title))
    cleaned_data$Year <- as.numeric(gsub(".xlsx", "", year_extract[2]))
    df <- rbind(df, cleaned_data)}
  return (df)
}
#Examples

# path = "~/Desktop/Projects/EDAV/edav-final-project-41/data/raw/demographics"
# title = 'Demographics_'
# function
# sheet_name = 'Ethnicity'
# sheet_name = 'Age Group'
# sheet_name = 'Religion'
# sheet_name = 'Education'
# sheet_name = 'Native Language'
# data_cleaner = clean_demographics

# path = "~/Desktop/Projects/EDAV/edav-final-project-41/data/raw/all_arrivals"
# title = 'Arrivals_'
# sheet_name = 'Detailed'
# data_cleaner = clean_arrival

# df_test3 <- combine_files(path, clean_demographics, sheet_name, title)
```

```{r}
# Cleaned all_arrivals data converted to csv files
save_cleaned_all_arrivals <- function(path) {
    read_path = paste(path, "/edav-final-project-41/data/raw/all_arrivals", sep='')
    write_path = paste(path, "/edav-final-project-41/data/clean", sep='')
    title = 'Arrivals_'
    sheet_name = 'Detailed'
    data_cleaner = clean_arriva
    arrival_df <- combine_files(read_path, data_cleaner, sheet_name, title)
    write.csv(arrival_df, file = paste(write_path, "all_arrivals.csv", sep=''), row.names = FALSE)
}

# Cleaned demographics data converted to csv files
save_cleaned_all_demographics <- function(path) {
  read_path = paste(path, "/edav-final-project-41/data/raw/demographics", sep='')
  write_path = paste(path, "/edav-final-project-41/data/clean", sep='')
  title = 'Demographics_'
  sheet_names = c('Ethnicity', 'Age Group', 'Religion', 'Education', 'Native Language')
  file_names = c('ethnicity', 'age_group', 'religion', 'education', 'native_language')
  data_cleaner = clean_demographics
  
  for (i in 1:length(sheet_names)) {
    demographic_df <- combine_files(read_path, data_cleaner, sheet_names[i], title)
    file_name = paste(write_path, file_names[i], sep='')
    write.csv(demographic_df, file = paste(file_name, '.csv', sep=''), row.names = FALSE)
  }
}
```

## Cleaned Data Format
```{r}
# Read cleaned data from github
arrival_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/all_arrivals.csv')
age_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/age_group.csv')
ethnicity_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/ethnicity.csv')
religion_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/religion.csv')
education_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/education.csv')
native_language_df = read.csv('https://raw.githubusercontent.com/tlzhu19/edav-final-project-41/master/data/clean/native_language.csv')
```

After cleaning the data, we have six '.csv' files that can be found [here](https://github.com/tlzhu19/edav-final-project-41/tree/master/data/clean). 

1. `all_arrivals.csv`: The total number of refugee resettlements to each of the 50 states in the US from 2009-2018. All raw files to make this file can be found [here](https://github.com/tlzhu19/edav-final-project-41/tree/master/data/raw/all_arrivals).
```{r}
kable(head(arrival_df))
```

2. `age_group.csv`: 
```{r}
kable(head(age_df))
```

3. `education.csv`: 
```{r}
kable(head(education_df))
```

4. `ethnicity.csv`: 
```{r}
kable(head(ethnicity_df))
```

5. `native_language.csv`: 
```{r}
kable(head(native_language_df))
```

6. `religion.csv`: 
```{r}
kable(head(religion_df))
```

# Missing Values
The datasets from RPC (Refugee Processing Center) did not contain any missing values. However, we also noticed that our data contained a single row called “Unknown State”. Since this would not be plotted in our maps, we decided that it would be better to remove the data. Additionally, when we converted the `State` column to factors, there were 56. The extra 6 states are:

* American Samoa
* District of Columbia
* Guam
* Puerto Rico
* Unknown State
* Virgin Islands

We removed these rows since we are just curious about the fifty states.


# Results


```{r}
# Compare ceilings with actual admissions
total_arrival_df <- arrival_df %>%
  group_by(Year) %>%
  summarise(Actual = sum(Inds))

# data from https://www.migrationpolicy.org/article/refugees-and-asylees-united-states#Refugee_Admission_Ceiling
Ceiling = c(80000, 80000, 80000, 76000, 70000, 70000, 70000, 85000, 50000, 45000)
# Actual = c(74654, 73311, 56424, 58238, 69926, 69987, 69933, 84995, 53716, 22491)
# Year = 2009:2018
# total_arrival_df = data.frame(ceiling, actual, year)

total_arrival_df$Ceiling = Ceiling

total_arrival_df <- total_arrival_df %>%
  gather(key, Individuals, Actual, Ceiling)

p <-ggplot(total_arrival_df, aes(x=Year, y=Individuals, colour=key)) +
      geom_line() +
      geom_point() +
      geom_vline(xintercept = 2016, color='gray') +
      annotate("text", x=2015, y =50000, label="Trump\nelected", color='gray', hjust=0) +
      ggtitle('Ceilings Vs Actual Refugee Resettlements')

ggplotly(p)
```


We take a closer look at the top 5 countries that have refugees resettled in the USA (https://data.newamericaneconomy.org/en/refugee-resettlement-us/).

```{r, fig.height=6}
# bar chart of the top countries faceted by year
# Todo: change to proportions?
top_countries <- age_df %>% 
  group_by(country, Year) %>%
  summarise(Inds = sum(Male) + sum(Female))

ggplot(top_countries, aes(x= reorder(country, -Inds), y=Inds)) + 
  geom_bar(stat="identity") +
  facet_grid(top_countries$Year) +
  labs(x = 'Country', y = 'Individuals') +
  ggtitle('')
```

The top 5 religions in the world are: Christianity, Islam, Hinduism, Buddhism, Sikhism (https://thecountriesof.com/top-5-largest-religions-in-the-world/). In Iraq, they separate the Muslim population into three categories: Muslim, Muslim Shiite, and Muslim Suni. For the purposes of this analysis, we will combine them as one.

```{r}
# religion_df2 <- religion_df %>% 
#   group_by(country, Year, Religion) %>%
#   summarise(Inds = sum(Male) + sum(Female)) %>%
#   mutate(Percent = Inds / sum(Inds))

# burma_religion_df = religion_df2[religion_df2$country == 'Burma', ]
# p <- ggplot(burma_religion_df, aes(x= reorder(Religion, -Percent), y=Percent)) + 
#   geom_bar(stat="identity") +
#   facet_grid(burma_religion_df$Year) +
#   labs(x = 'Religion', y = 'Individuals') +
#   ggtitle('Religion (Burma)')


muslim_names = c('Moslem', 'Moslem Suni', 'Moslem Shiite')

religion_df3 <- religion_df %>% 
  group_by(Year, Religion) %>%
  summarise(Inds = sum(Male) + sum(Female)) %>%
  mutate(Percent = Inds / sum(Inds))

# rename to Moslem
religion_df3[religion_df3$Religion %in% muslim_names, 'Religion'] <- 'Moslem'

innerFunc <- function(sub_df){
  row <- head(sub_df, 1)
  row$Inds <- sum(head(sub_df$Inds))
  return(row)
}

religions = c('Christian', 'Moslem', 'Buddhist', 'Hindu')
subset_religion_df3 <- religion_df3[religion_df3$Religion %in% religions, ]

# combines Moslem for each year
create_religion_df <- function(current_df) {
  df <- do.call(rbind,
                  by(data=current_df[current_df$Year==2009, ],
                     INDICES=current_df[current_df$Year==2009, ]$Religion, FUN=innerFunc))
    
  for (year in c(2010:2018)) {
    df2 <- do.call(rbind,
                   by(data=current_df[current_df$Year==year, ],
                     INDICES=current_df[current_df$Year==year, ]$Religion, FUN=innerFunc))
    df <- rbind(df, df2)
}
  
  return(df)
}

subset_religion_df4 <- create_religion_df(subset_religion_df3)


p2 <- ggplot(subset_religion_df4, aes(x= reorder(Religion, -Percent), y=Percent)) + 
  geom_bar(stat="identity") +
  facet_grid(subset_religion_df4$Year) +
  labs(x = 'Religion', y = 'Individuals Percentages') +
  ggtitle('Religion Over Time')

p2
```

```{r}
# Age group
age_df2 <- age_df %>% 
  group_by(Year, Age.Group) %>%
  summarise(Inds = sum(Male) + sum(Female)) %>%
  mutate(Percent = Inds / sum(Inds))

p3 <- ggplot(age_df2, aes(x= reorder(Age.Group, -Percent), y=Percent)) + 
  geom_bar(stat="identity") +
  facet_grid(age_df2$Year) +
  labs(x = 'Age Group', y = 'Individuals Percentages') +
  ggtitle('Age Group Over Time')

p3
```

```{r}
total_demo <- ethnicity_df
total_demo$Total <- as.numeric(total_demo$Total)
change_in_pop <- total_demo %>%
  group_by(Year, country) %>%
  summarize(sum=sum(Total))
ggplot(change_in_pop, aes(Year, sum, color=country))+
  geom_line() + 
  xlab("Year") + 
  ylab("Number of Individuals") +
  labs(title = "Change in Refugee Resettlement to the US 
       of Top 5 Refugee Origins from 2009-2018 ",
       caption = "Figure: Times series graph of top 5 refugee origins with the most number refugee resettlements.") +
  theme(legend.title = element_blank())
```


```{r, fig.width=8}
df_map <- arrival_df

# remove extra 6 'States'
df_map <- df_map[!df_map$State %in% c('American Samoa', 'District of Columbia', 'Guam', 'Puerto Rico', 'Unknown State', 'Virgin Islands'), ]
df_map$State <- as.factor(df_map$State)
df_map$code <- state.abb[match(df_map$State, state.name)]

#add Wyoming and Montana because they are misisng 
df_add <- data.frame(State = c('Wyoming','Wyoming','Montana'), Cases = c(0, 0, 0), Inds = c(0, 0, 0), Year = c(2009, 2010, 2010), code = c('WY', 'WY', 'MT'))

df_map <- rbind(df_map, df_add)

df_map$State <- tolower(df_map$State)

states <- map_data('state')

# fill = Inds? or cases?
map_basic <- ggplot(df_map, aes(map_id = State)) + 
  geom_map(aes(fill = Inds), map = states) + 
  expand_limits(x = states$long, y = states$lat) +
  facet_wrap(~Year) +
  scale_fill_gradient(low="#D6EAF8", high="#154360") +
  ggtitle('Refugee Resettlement in the US by State')

map_basic
```


# Interactive Component
```{r}
# remove extra 6 'States'
arrival_df2 <- arrival_df[!arrival_df$State %in% c('American Samoa', 'District of Columbia', 'Guam', 'Puerto Rico', 'Unknown State', 'Virgin Islands'), ]
arrival_df2$State <- as.factor(arrival_df2$State)
arrival_df2$code <- state.abb[match(arrival_df2$State, state.name)]

gg <- ggplot(arrival_df2, aes(code, Inds, frame = Year)) + 
  geom_bar(stat="identity", position='identity') +
  labs(x = "State", y="Individuals") +
  ggtitle('Refugee Resettlements By State from 2009-2018') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
ggplotly(gg)
```

```{r}
library(maps)
arrival_df2$hover <- with(arrival_df2, paste(State, '<br>', "Inds", Inds))

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

df_2009 <- arrival_df2[arrival_df2$Year==2009, ]
p2 <- plot_geo(df_2009, locationmode = 'USA-states') %>%
  add_trace(
    z = ~Inds,
    locations = ~code,
    text = ~hover,
    color = ~Inds,
    colors = 'Blues'
  ) %>%
  colorbar(title = "Number of Individuals") %>%
  layout(
    title = '2009 Refugee Resettlements in USA',
    geo = g
  )

p2
```

# Conclusion

